---
title: "DEMO: Bivariate Choropleth in Leaflet using R"
author: "Christopher Belanger, PhD"
date: "July 1,2021"
output: html_document
---
  
  

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(sf)
library(leaflet)

source("../R/bivariate_choropleth.R")

ons_shp <- onsr::get_ons_shp()
ons_table_auto <- read_csv("../../cpso-targets/outputs/neighbourhood_tables/ons_table_auto_rurality.csv")

ons_data <- onsr::get_ons_data()

drive_time_fr <- ons_table_auto %>%
  select(ONS_ID,
         time_fr) %>%
  mutate(time_fr = round(time_fr, digits = 2))

fr_only <- ons_data %>%
  filter(polygon_attribute == "CDP269b") %>%
  select(ONS_ID,
         fr_only_pct = value) %>%
  distinct() %>%
  mutate(ONS_ID = as.numeric(ONS_ID)) %>%
  mutate(fr_only_pct = round(fr_only_pct, digits = 2))

data_forplot <- ons_shp %>%
  select(ONS_ID, Name) %>%
  left_join(drive_time_fr, by = "ONS_ID") %>%
  left_join(fr_only, by = "ONS_ID")
# 
# 
# 
# renfrew_csds <- sf::read_sf("shapefiles/renfrew_csds.shp") %>%
#   select(CSDUID, CSDNAME)
# 
# csd_median_household_income <- read_csv("data/renfrew_csd_census_2016.csv", col_types = cols(.default = "c")) %>%
#   filter(TEXT_ID == "13001") %>%
#   select(CSDUID = GEO_ID,
#          CSDNAME = GEO_NAME_NOM,
#          median_total_household_income = T_DATA_DONNEE) %>%
#   mutate(median_total_household_income = as.numeric(median_total_household_income))
# 
# csd_fp_popwt_avg <- read_csv("data/csd_fp_popwt_avg.csv", col_types = cols(.default = "c")) %>%
#   mutate(avg_mins_popwt = as.numeric(avg_mins_popwt))
# 
# # join it all together
#  renfrew_csds <- renfrew_csds %>%
#   left_join(csd_fp_popwt_avg, by = c("CSDUID", "CSDNAME")) %>%
#   left_join(csd_median_household_income, by = c("CSDUID", "CSDNAME"))

```



The map below is a *bivariate choropleth* which shows the relation between two variables using colour. In this case we're mapping Ottawa's neighbourhoods based on their travel time to primary care (from a forthcoming study) and their percentage of French-only-speaking residents (from the 2016 census).

This is a demo of the function `addBivariateChoropleth()`, which is written in R, CSS, and HTML, and extends the existing `leaflet::addPolygons()` function. As far as I know this is a novel contribution, since I wasn't able to find any leaflet implementation of bivariate choropleths.

*Please note this is a work in progress.*

```{r}

leaflet() %>% 
  addProviderTiles(provider =  "CartoDB.Positron") %>%
  addBivariateChoropleth(data_forplot, 
                         time_fr, 
                         fr_only_pct,
                         var1_label = "Drive Time to FR-Speaking PCP",
                         var2_label = "% FR-Only Speakers",
                         region_names = "Name",
                         weight = 1,
                         fillOpacity = 0.7,
                         color = "grey",
                         highlightOptions = leaflet::highlightOptions(color = "orange",
                                                                      weight = 2,
                                                                      opacity = 1))
```


Regions that are darker pink have higher percentages of French-only speaking residents; regions that are more blue have higher travel times to French-speaking primary care. 


